---
title: "gs-model-initial"
author: "Gabriel De La Rosa"
date: "5/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(growthrates)
library(tidyverse)
```

Let's begin exploring Chris's simple fisheries model. Harvest is given as...

h<sub>t</sub> = qE<sub>t</sub>x<sub>t</sub>

where...

* x<sub>t</sub> = population size
* h<sub>t</sub> = harvest,
* E<sub>t</sub> = fishing effort
* q = a constant

profit is...

profit = p * h<sub>t</sub>

where p is the price

And stock growth as...
\[x_{t+1} = x_{t} + rx_{t} (1 - \frac{x_{t}}{K}) - h_{t}\]


where...

* r = intrinsic growth rate
* K = carrying capacity


```{r}

fish_model <- function(years, r, K, q, p, E) {
  # Set original stock size to 1...
  stock_size <- vector(mode = "numeric", length = years)
  stock_size[1] <- 1
  
  # Set original harvest to 0...
  harvest <- vector(mode = "numeric", length = years)
  harvest[1] <- 0
  
  # Make a profit vector...
  profit <- vector(mode = "numeric", length = years)
  profit[1] <- 0
  
  # Create an iterative for loop that'll sequentially run this...
  for(iYear in 2:years) {
  stock_size[iYear] <- stock_size[iYear-1] + (r * stock_size[iYear-1]) * (1 - (stock_size[iYear-1]/K)) - q * E * stock_size[iYear-1]
  }
  
  for(iYear in 1:years){
  # Add in yearly harvest:
  harvest[iYear] <- q * E * stock_size[iYear]
  
  # And yearly profit:
  profit[iYear] <- p * harvest[iYear]
  }
  
  # And make a data frame from both of them.
  data.frame(year = seq(1, years, by = 1), 
             stock_size = stock_size, 
             harvest = harvest, 
             profit =profit)
}
```

``` {r}

# Then, we can run it with a fishing effort of 0...
initial_run <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 0)

# And compare it to a fishing effort of 10
second_run <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 10)

# Plotting them both shows us nice things
# First run:
ggplot(initial_run) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# Second run:
ggplot(second_run) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

```

Can we maximize the total profit? Let's play around with a few extremes.

```{r}
# Fishing effort of 99...
max_run <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 99)

ggplot(max_run) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# What about a 50 fishing effort?
run_50 <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 50)

ggplot(run_50) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# Whoa! Big jump
run_20 <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 20)

ggplot(run_20) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# This is still going up...
run_25 <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 25)
ggplot(run_20) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# Thirty is the highest!
run_30 <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 30)
ggplot(run_30) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")

# And this starts dropping again
run_40 <- fish_model(years = 100, r = 0.6, K = 100, q = .01, p = 10, E = 40)
ggplot(run_40) +
  geom_line(aes(x = year, y = stock_size),
            color = "blue") +
  geom_line(aes(x = year, 
                y = harvest),
                color = "red") +
  geom_line(aes(x = year, y = profit),
            color = "green")
```


### Thinking about maximizing total profit: 

Can we solve for max profit mathematically?
